---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
data = read.csv("activity.csv")
```

```{r, eval=FALSE}
library("qpcR")
```

```{r, include=FALSE}
library("qpcR")
```

## What is mean total number of steps taken per day?

```{r fig.width=7, fig.height=6}
steps<-aggregate(data$steps,by=list(Category=data$date), FUN=sum)
hist(steps$x, main="Total number of steps taken each day", 
     xlab="Steps taken each day", 
     border="blue", 
     col="green",
     las=1, 
     breaks=5)
```

```{r}
str(steps)
summary(steps)
```

The mean of the total number of steps taken per day is: 
```{r}
as.integer(mean(steps$x, na.rm = TRUE))
```

The median of the total number of steps taken per day is: 
```{r}
as.integer(median(steps$x, na.rm = TRUE))
```


## What is the average daily activity pattern?

```{r}
times<-rowsum(data$steps,group = data$interval, na.rm = TRUE)
times<-as.integer(times/30)
times<-qpcR:::cbind.na(data$interval,times)
times<-na.omit(times)
colnames(times)<-c("interval","average")
times<-data.frame(times)
```

```{r fig.width=7, fig.height=6}
plot(times$interval,times$average,type="l", main="Average number of steps taken",
     xlab="5-minute interval",
     ylab="Average number of steps")
```

The maximum number of steps is contained in the 
```{r include=TRUE}
times[which.max(times$average),][[1]]
```
5-minute interval.

## Imputing missing values
The total number of missing values is:
```{r}
sum(is.na(data$steps))
```

```{r}
data_new<-data


for(i in 1:dim(data_new)[1])
{
    
    if (is.na(data_new[i,1]))
    {
        
        temp2<-which(times$interval==data_new[i,3])
        
        data_new[i,1]<-times[temp2,2]
        
    }
    
}

steps_new<-aggregate(data_new$steps,by=list(Category=data_new$date), FUN=sum)
```

```{r fig.width=7, fig.height=6}
hist(steps_new$x, main="Total number of steps taken each day - with estimates for NAs", 
     xlab="Steps taken each day", 
     border="blue", 
     col="green",
     las=1, 
     breaks=5)
```


```{r}
str(steps_new)
summary(steps_new)
```

The mean of the total number of steps taken per day is: 
```{r}
as.integer(mean(steps_new$x, na.rm = TRUE))
```

The median of the total number of steps taken per day is: 
```{r}
as.integer(median(steps_new$x, na.rm = TRUE))
```
Filling the missing values has raised the number of steps, the mean and the medial of the total number of steps taken per day.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
data_new$day<-as.numeric(strftime(as.Date(data_new$date, "%Y-%m-%d"), "%u"))
data_new$day<-replace(data_new$day,data_new$day<7,0)
data_new$day<-replace(data_new$day,data_new$day==7,1)
data_new$day<-factor(data_new$day,labels = c("weekday","weekend"))


times_weekend<-rowsum(subset(data_new,data_new$day=='weekend')$steps,group = subset(data_new,data_new$day=='weekend')$interval, na.rm = TRUE)
times_weekend<-as.integer(times_weekend/30)
times_weekend<-qpcR:::cbind.na(data$interval,times_weekend)
times_weekend<-na.omit(times_weekend)
colnames(times_weekend)<-c("interval","average")
times_weekend<-data.frame(times_weekend)

times_weekday<-rowsum(subset(data_new,data_new$day=='weekday')$steps,group = subset(data_new,data_new$day=='weekday')$interval, na.rm = TRUE)
times_weekday<-as.integer(times_weekday/30)
times_weekday<-qpcR:::cbind.na(data$interval,times_weekday)
times_weekday<-na.omit(times_weekday)
colnames(times_weekday)<-c("interval","average")
times_weekday<-data.frame(times_weekday)
```

```{r fig.width=7, fig.height=10}
par(mfrow=c(2,1))
plot(times_weekday$interval,times_weekday$average,type="l", xlab="5-minute interval", ylab="Weekdays")
title("Average number of steps taken")
plot(times_weekend$interval,times_weekend$average,type="l", xlab="5-minute interval", ylab="Weekends")
```

From the plot we can see that while at weekdays most steps are centered between the 800th and 1000th 5-minute interval, at weekends the steps are more scattered between the 800th and 2000th 5-minute interval.